\Large Last Writer Slices and Communication Traps

### 2013-11-07 paper.tex
committing?
committing before changing networks

Data provenance information is directly useful during program debugging.
Work on debugging~\cite{whylinechi,whylineicse} in the HCI and Software
Engineering community shows that when programmers debug programs, they benefit
from the ability to directly ask such provenance (``why'') questions of their
debugger.

@@

First, many of these techniques track only an {\em ad hoc} subset of
dependences.  For example, DefUse~\cite{defuse} tracks inter-thread WAR
dependences, but not WAW dependences.  Recon tracks some inter-thread WAW and
RAW dependences and approximates WAR dependences.  These techniques are useful
despite an incomplete set of dependences.  Second, many of these systems rely
on statistical models of software
behavior~\cite{avio,cci,defuse,recon,bugaboo,falcon,dmtracker,aviso,threadcriticality,schedpredictionmodel}.
Models often tolerate noise, so many of these techniques can relax the
correctness constraints of their implementation and remain useful.  For
example, some systems have chosen not to enforce atomicity of each memory
access and its dependence tracking code, compromising precision for a boost in
performance.

@@

{\bf TODO: Figure?}

@@

\ctraps delivers software traps before all read and write operations executed
by one thread that access a memory location that was last written by another
thread, {\em i.e.}, communicating memory operations.  Traps are simply function
calls made before such operations execution.  \ctraps implements traps using a
combination of compiler and runtime library support.

@@

We built our runtime system from scratch in about 1000 lines of C and C++ code.
We implemented the LWT as a fixed size 2GB array of 64-bit words.  Each entry
packs a program point and a thread identifier into the 64-bit word.  By
default, program points are single instruction addresses, but we also support
limited (2-address) context-sensitivity by packing the current instruction's
address and the nearest return address into a 64-bit LWT entry. When an access
to a memory location occurs, we index into the LWT with the lower bits of the
location's address.  Our prototype implementation uses a lossy resolution
policy for hash collisions.  We use this policy in our prototype because it is
faster and simpler than a chaining concurrent hash table implementation.
Collisions are extremely rare, so this simplification is unlikely to be a
problem even in a more industrial strength implementation.

@@

\subsection{Context-Sensitivity}
{\bf TODO: Was this LWS or CTRAPS-NOOP or what?}
We briefly evaluated the cost of context-sensitivity for \ctraps.  With a
bounded 2-address context, the average overhead for server programs was 18\%
and the average overhead for PARSEC was 145\%.  These data show that a
restricted form of context-sensitive analysis is possible in \ctraps with
overheads low enough for production.

@@

@@

\section{Conclusions}
This work stems observes that we can collect a useful form of data provenance
information with overheads low enough for use in deployed systems.

>>>>

Data provenance information is directly useful during program debugging.  Work
on debugging~\cite{whylinechi,whylineicse} in the HCI and Software Engineering
community shows that when programmers debug programs, they benefit from the
ability to directly ask such provenance (``why'') questions of their debugger.
Note that that like prior work on debugging with provenance, in this work we
focus on {\em debugging} -- understanding the root cause of a failure -- rather than {\em bug detection} -- finding new, unknown bugs.

@@

First, many of these techniques track only an {\em ad hoc} subset of
dependences.  For example, DefUse~\cite{defuse} tracks inter-thread WAR
dependences, but not WAW dependences.  Recon~\cite{recon} tracks some inter-thread WAW and
RAW dependences and approximates WAR dependences.  These techniques are useful
despite an incomplete set of dependences.  Second, many of these systems rely
on statistical models of software
behavior~\cite{avio,cci,defuse,recon,bugaboo,falcon,dmtracker,aviso,threadcriticality,schedpredictionmodel}.
Models often tolerate noise, so many of these techniques can relax the
correctness constraints of their implementation and remain useful.  For
example, some systems have chosen not to enforce atomicity of each memory
access and its dependence tracking code, compromising precision for a boost in
performance.

@@

%{\bf TODO: Figure?}

@@

\ctraps delivers software traps before all read and write operations executed
by one thread that access a memory location that was last written by another
thread, {\em i.e.}, communicating memory operations.  Traps are simply function
calls made before such operations execution.  \ctraps implements traps using a
combination of compiler and runtime library support.  Note that we use the name
``trap'' for the interposition points in \ctraps, but we implement traps using 
function calls only, as opposed to signals or some other mechanism.

@@

We built our runtime system from scratch.  We implemented the LWT as a fixed
size 2GB array of 64-bit words, fixing our memory overheads.  Each entry packs a program point and a thread
identifier into the 64-bit word.  By default, program points are single
instruction addresses, but we also support limited (2-address)
context-sensitivity by packing the current instruction's address and the
nearest return address into a 64-bit LWT entry. When an access to a memory
location occurs, we index into the LWT with the lower bits of the location's
address.  Our prototype implementation uses a lossy resolution policy for hash
collisions.  We use this policy in our prototype because it is faster and
simpler than a chaining concurrent hash table implementation.  Collisions are
extremely rare, so this simplification is unlikely to be a problem even in a
more industrial strength implementation.

@@

\subsection{Context-Sensitivity}
%{\bf TODO: Was this LWS or CTRAPS-NOOP or what?}
We briefly evaluated the cost of context-sensitivity using \ctraps with no-op
trap handlers.  With a bounded 2-address context, the average overhead for
server programs was 18\% and the average overhead for PARSEC was 145\%.  These
data show that a restricted form of context-sensitive analysis is possible in
\ctraps with overheads low enough for production.

@@

Bad Apple Tracking~\cite{bondapple} is also a form of debugging with data
provenance information. This work records the {\em origin} of an unusable value
and can report it to a programmer when it is used in a way that causes a
failure.

@@

\section{Conclusions}
This work observes that we can collect a useful form of data provenance
information with overheads low enough for use in deployed systems.
